---
title: "02_SeuratSetup_SplitCellLines"
author: "Andrea Gardner"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(Seurat)
`%notin%` <- negate(`%in%`)
```

# ClonMapper Duo: Seurat Object setup

## Introduction
This notebook separates each cell line into its own Seurat object and performs normalization on each. Cell cycle regression is performed to optimize clustering and UMAP visualization.
```{r}
dat_qc3_scdbl <- readRDS('./data/rds/seurat_obj_dat_qc3_scdbl.rds')
# remove doublets
dat_qc3_scdbl_singles <- subset(dat_qc3_scdbl, subset = scDblFinder.class == 'singlet')
```

```{r}
# subset data by cell line
dat_h1806 <- subset(dat_qc3_scdbl_singles, subset = cellLine == 'H1806')
dat_mb231 <- subset(dat_qc3_scdbl_singles, subset = cellLine == 'MB231')
```

```{r}
# check QC metrics on subsets
VlnPlot(dat_h1806, pt.size = 0.10, alpha=0.05, features = c("nFeature_RNA", "nCount_RNA", "percent_mt"), ncol = 3, group.by = 'condition')
VlnPlot(dat_mb231, pt.size = 0.10, alpha=0.05, features = c("nFeature_RNA", "nCount_RNA", "percent_mt"), ncol = 3, group.by = 'condition')
```


# log normalize data
```{r}
dat_h1806_norm <- NormalizeData(dat_h1806, normalization.method = "LogNormalize", scale.factor = 10000)
dat_mb231_norm <- NormalizeData(dat_mb231, normalization.method = "LogNormalize", scale.factor = 10000)
```

## Find Highly variable genes and scale data
```{r}
# find HVGs
dat_h1806_norm <- FindVariableFeatures(dat_h1806_norm, selection.method = "vst", nfeatures = 10000)
dat_mb231_norm <- FindVariableFeatures(dat_mb231_norm, selection.method = "vst", nfeatures = 10000)

# scale data
dat_h1806_norm <- ScaleData(dat_h1806_norm)
dat_mb231_norm <- ScaleData(dat_mb231_norm)
```

# Run PCA
```{r}
dat_h1806_norm <- RunPCA(dat_h1806_norm)
print(dat_h1806_norm[["pca"]], dims = 1:5, nfeatures = 5)
VizDimLoadings(dat_h1806_norm, dims = 1:2, reduction = "pca")
DimPlot(dat_h1806_norm, reduction = "pca", group.by = "condition") + scale_color_manual(values=c('dodgerblue','goldenrod','black','grey80')) + ggtitle('HCC1806')
ElbowPlot(dat_h1806_norm, ndims = 50) + ggtitle('HCC1806')

dat_mb231_norm <- RunPCA(dat_mb231_norm)
print(dat_mb231_norm[["pca"]], dims = 1:5, nfeatures = 5)
VizDimLoadings(dat_mb231_norm, dims = 1:2, reduction = "pca")
DimPlot(dat_mb231_norm, reduction = "pca", group.by = "condition") + scale_color_manual(values=c('dodgerblue','goldenrod','black','grey80')) + ggtitle('MDA-MB-231')
ElbowPlot(dat_mb231_norm, ndims = 50) + ggtitle('MDA-MB-231')
```

# Check cell cycle effects
```{r}
# cell cycle genes come pre-loaded with Seurat
s.genes <- cc.genes$s.genes
g2m.genes <- cc.genes$g2m.genes

# assign cell cycle scores
dat_h1806_norm <- CellCycleScoring(dat_h1806_norm, s.features = s.genes, g2m.features = g2m.genes, set.ident = TRUE)
dat_mb231_norm <- CellCycleScoring(dat_mb231_norm, s.features = s.genes, g2m.features = g2m.genes, set.ident = TRUE)

# visualize cell cycle on umap}
DimPlot(dat_h1806_norm, reduction = "pca", group.by = "Phase") 
DimPlot(dat_mb231_norm, reduction = "pca", group.by = "Phase") 
```

# Save data before cell cycle regression
```{r}
saveRDS(dat_mb231_norm, './data/rds/mb231_singlets.rds')
saveRDS(dat_h1806_norm, './data/rds/h1806_singlets.rds')
```

# Regress cell cycle
This step can take a very long time (6+ hours), recommend running as script.
```{r}
# dat_mb231_norm.cc <- ScaleData(dat_mb231_norm, vars.to.regress = c("S.Score", "G2M.Score"), features = rownames(dat_mb231_norm))
# saveRDS(dat_mb231_norm.cc, paste0('./data/data/rds/mb231_ccregress_singlets.rds'))
# 
# dat_h1806_norm.cc <- ScaleData(dat_h1806_norm, vars.to.regress = c("S.Score", "G2M.Score"), features = rownames(dat_h1806_norm))
# saveRDS(dat_h1806_norm.cc, paste0('./data/data/rds/h1806_ccregress_singlets.rds'))
```


## Run PCA and UMAP on cc regressed data
Load cell cycle regressed data
```{r}
dat_mb231_cc <- readRDS('./data/rds/mb231_ccregress_singlets.rds')
dat_h1806_cc <- readRDS('./data/rds/h1806_ccregress_singlets.rds')
```



### PCA
Run PCA and check that regressed data does not separated based on cell cycle phase
```{r}
dat_h1806_cc_red <- RunPCA(dat_h1806_cc)
ElbowPlot(dat_h1806_cc_red, ndims = 50) 
DimPlot(dat_h1806_cc_red, reduction = "pca", group.by = 'Phase')

dat_mb231_cc_red <- RunPCA(dat_mb231_cc)
ElbowPlot(dat_mb231_cc_red, ndims = 50) 
DimPlot(dat_mb231_cc_red, reduction = "pca", group.by = 'Phase')
```

### UMAP and clustering
Will now compute UMAP and run Louvain clustering. Will analyze clustering decisions after barcode assignment in future notebooks under the assumption that most clones will primarily be in one transcriptomic cluster
```{r}
dat_h1806_cc_red <- FindNeighbors(dat_h1806_cc_red, dims = 1:20)
dat_h1806_cc_red <- RunUMAP(dat_h1806_cc_red, dims = 1:15, min.dist = 0.4) 

DimPlot(dat_h1806_cc_red, reduction = "umap") 
DimPlot(dat_h1806_cc_red, reduction = "umap", group.by = "condition") + scale_color_manual(values=c('dodgerblue','goldenrod','black','grey80'))

dat_h1806_cc_red <- FindClusters(dat_h1806_cc_red, resolution = 0.18)
DimPlot(dat_h1806_cc_red, reduction = "umap")
```

```{r}
dat_mb231_cc_red <- FindNeighbors(dat_mb231_cc_red, dims = 1:20)
dat_mb231_cc_red <- RunUMAP(dat_mb231_cc_red, dims = 1:15, min.dist = 0.4)

DimPlot(dat_mb231_cc_red, reduction = "umap") 
DimPlot(dat_mb231_cc_red, reduction = "umap", group.by = "condition") + scale_color_manual(values=c('dodgerblue','goldenrod','black','grey80'))

dat_mb231_cc_red <- FindClusters(dat_mb231_cc_red, resolution = 0.18)
DimPlot(dat_mb231_cc_red, reduction = "umap")
```

## Save cc regressed and clustered objects
Save regressed and Leiden clustered data. Note that cc regressed Seurat objects are large (~10 GB) and takes an excessive amount of time to save ~10 minutes per object
```{r}
saveRDS(dat_mb231_cc_red, './data/rds/mb231_leiden_ccregres_singlets.rds')
saveRDS(dat_h1806_cc_red, './data/rds/h1806_leiden_ccregres_singlets.rds')
```

As we want to use the UMAP visualization and cell clustering from the cell cycle regressed data, but want to perform downstream analysis on the non-regressed data, we will take the calculated UMAP and clustering from the cc regressed object and append it to the non-cc regressed object
```{r}
# load data pre-cc regression if not already loaded
dat_mb231_norm <- readRDS('./data/rds/mb231_singlets.rds')
dat_h1806_norm <- readRDS('./data/rds/h1806_singlets.rds')

# assign umap reduction from cc regressed data to pre-cc regressed object
dat_mb231_norm@reductions$umap <- dat_mb231_cc_red@reductions$umap
dat_h1806_norm@reductions$umap <- dat_h1806_cc_red@reductions$umap

# get clustering meta data from cc regressed object and add to pre-cc regressed object
dat_mb231_norm@meta.data <- left_join(dat_mb231_norm@meta.data %>% rownames_to_column('cellid'), dat_mb231_cc_red@meta.data %>% rownames_to_column('cellid') %>% select(cellid, seurat_clusters), by='cellid') %>% column_to_rownames('cellid')
dat_h1806_norm@meta.data <- left_join(dat_h1806_norm@meta.data  %>% rownames_to_column('cellid'), dat_h1806_cc_red@meta.data %>% rownames_to_column('cellid') %>% select(cellid, seurat_clusters), by='cellid') %>% column_to_rownames('cellid')

# update sample identities to cluster
Idents(dat_mb231_norm) <- 'seurat_clusters'
Idents(dat_h1806_norm) <- 'seurat_clusters'

# check that reductions and clustering was added
DimPlot(dat_mb231_norm, reduction = 'umap')
DimPlot(dat_h1806_norm, reduction = 'umap')
```

## Save pre-cc but cc-clustered data
```{r}
saveRDS(dat_mb231_norm, './data/rds/mb231_ccclustered_precc_singlets.rds')
saveRDS(dat_h1806_norm, './data/rds/h1806_ccclustered_precc_singlets.rds')
```






