```{r}
suppressWarnings(library(Seurat))
suppressWarnings(library(tidyverse))
suppressWarnings(library(igraph))
suppressWarnings(library(scales))
source('functions/networkPlots.R')
source('functions/color_mapping.R')
`%notin%` <- negate(`%in%`)
fig_path <- "./plots/network/"
```


## Load clustered data with assigned barcode ids
```{r}
mb231 <- readRDS('./data/rds/mb231_hclust_nb-barassigned_ccclustered_precc_singlets.rds')
h1806 <- readRDS('./data/rds/h1806_hclust_nb-barassigned_ccclustered_precc_singlets.rds')

h1806@meta.data$replicate[is.na(h1806@meta.data$replicate)==TRUE] <- '0'
mb231@meta.data$replicate[is.na(mb231@meta.data$replicate)==TRUE] <- '0'
```

## Choose the minimum number of fusion cells represented per clone per replicate required per replicate for classification
```{r}
min_fusion_clone_cells_per_rep = 5
```


## Create network representation of fusion events in the population
```{r, warning=FALSE, quiet=TRUE}
net_h1806 <- layout_network(obj = h1806, 
                       cellline = 'h1806', 
      min_cell_per_fusion_clone = min_fusion_clone_cells_per_rep, 
             min_cell_per_clone = 1, 
                    layout_type = 'nicely', 
                   plot_scaling = 10)

net_mb231 <- layout_network(obj = mb231, 
                       cellline = 'mb231', 
      min_cell_per_fusion_clone = min_fusion_clone_cells_per_rep, 
             min_cell_per_clone = 1, 
                    layout_type = 'nicely', 
                   plot_scaling = 10)

h1806_plots <- make_all_plots(net_h1806[[1]], cellline='HCC1806'   , scale='none')
mb231_plots <- make_all_plots(net_mb231[[1]], cellline='MDA-MB-231', scale='none')
 
h1806_plots
mb231_plots
```

# Save figures
```{r}
width = 5
height = width
res = 600

cellline = 'h1806'
SaveFigure(h1806_plots[[1]], paste0(cellline,'_constellation')      , type = "png", width, height, res)
SaveFigure(h1806_plots[[2]], paste0(cellline,'_network_replicate_1'), type = "png", width, height, res)
SaveFigure(h1806_plots[[3]], paste0(cellline,'_network_replicate_2'), type = "png", width, height, res)
SaveFigure(h1806_plots[[4]], paste0(cellline,'_network_replicate_3'), type = "png", width, height, res)
SaveFigure(h1806_plots[[5]], paste0(cellline,'_network_replicate_4'), type = "png", width, height, res)
SaveFigure(h1806_plots[[6]], paste0(cellline,'_network_all')        , type = "png", width, height, res)

cellline = 'mb231'
SaveFigure(mb231_plots[[1]], paste0(cellline,'_constellation')      , type = "png", width, height, res)
SaveFigure(mb231_plots[[2]], paste0(cellline,'_network_replicate_1'), type = "png", width, height, res)
SaveFigure(mb231_plots[[3]], paste0(cellline,'_network_replicate_2'), type = "png", width, height, res)
SaveFigure(mb231_plots[[4]], paste0(cellline,'_network_replicate_3'), type = "png", width, height, res)
SaveFigure(mb231_plots[[5]], paste0(cellline,'_network_replicate_4'), type = "png", width, height, res)
SaveFigure(mb231_plots[[6]], paste0(cellline,'_network_all')        , type = "png", width, height, res)
```


# Classify fusion cells and the number of fusion events each fusion parent is invovled in and add this to meta data
```{r}
h1806_classified <- classify_cells_fusion_and_parent(h1806, min_fusion_clone_cells_per_rep = min_fusion_clone_cells_per_rep)
mb231_classified <- classify_cells_fusion_and_parent(mb231, min_fusion_clone_cells_per_rep = min_fusion_clone_cells_per_rep)
```

# Save classifed Seurat objects
```{r}
# saveRDS(mb231_classified, './data/rds/mb231_cellclass_hclust_nb-barassigned_ccclustered_precc_singlets.rds')
# saveRDS(h1806_classified, './data/rds/h1806_cellclass_hclust_nb-barassigned_ccclustered_precc_singlets.rds')
```

# Plot number of fusion partners
```{r}
source('functions/n_fusion_partners_plots.R')

partners_summary_plot <- n_fusion_partners_plot(h1806_classified, mb231_classified, 
                                                x.text.size = 10, 
                                                y.text.size=10, 
                                                label.font.size=3)

SaveFigure(partners_summary_plot, paste0('n_fusion_partners_plot'), type = "png", width=7, height=4, res=600)
```


```{r}
source('functions/n_fusion_partners_plots.R')
mb_binom <- binom_test_n_partners(mb231_classified, pcut = 0.05, label.font.size=3.5, x.text.size=12, y.text.size=12, vertical=FALSE)
hc_binom <- binom_test_n_partners(h1806_classified, pcut = 0.05, label.font.size=3.5, x.text.size=12, y.text.size=12, vertical=FALSE)


mb_binom[[1]]
hc_binom[[1]]
SaveFigure(mb_binom[[1]], paste0('n_fusion_partners_plot_binom_mb231'), type = "png", width=5, height=6, res=600)
SaveFigure(hc_binom[[1]], paste0('n_fusion_partners_plot_binom_h1806'), type = "png", width=5, height=6, res=600)
```

```{r}
sessionInfo()
```
