```{r, show=FALSE, warning=FALSE, quiet=TRUE}
library(tidyverse)
library(Seurat)
library(cowplot)
`%notin%` <- negate(`%in%`)
source('functions/color_mapping.R')
source('functions/ParentProgenyUMAPs_DE_functions.R')
source('functions/ggumap.R')
fig_path <- "./plots/parent_progeny_eachClone_analysis/"
```

```{r load data}
mb231 <- readRDS('./data/rds/mb231_colors_cellclass_hclust_nb-barassigned_ccclustered_precc_singlets.rds')
mb231_markers <- read.csv('outdata/clusters/metacluster_markergenes_mb231.tsv',sep='\t')

h1806 <- readRDS('./data/rds/h1806_colors_cellclass_hclust_nb-barassigned_ccclustered_precc_singlets.rds')
h1806_markers <- read.csv('outdata/clusters/metacluster_markergenes_h1806.tsv',sep='\t')
```

For each parent-parent-fusion set, want to identify marker genes of each parent and then marker genes of fusion clones versus parent
Will define a marker gene as something that is present in >= 75% of cells within that clonal lineage
Will only look at clones with at least `cell_min` cells per population in a fusion set
```{r}
cellline='mb231'

if(cellline=='h1806'){
  obj <- h1806
  markers_full <- h1806_markers
}

if(cellline=='mb231'){
  obj <- mb231
  markers_full <- mb231_markers
}


#### Create seurat object subset containing only fusion trios (parent-parent-fusion) with at least cell_min cells each
obj_trio_min <- subset_fusion_trios_min_cell(obj, parent_conds = c('initial'), cell_min=20)

# scale data on all features
obj_trio_min@assays$RNA$scale.data <- NULL 
obj_trio_min <- ScaleData(obj_trio_min, features = rownames(obj_trio_min))

fusepairs <- obj_trio_min@meta.data %>%
  filter(fusion=='fusion') %>%
  select(gbarid, rbarid, barpairid) %>% 
  arrange(gbarid) %>% 
  distinct() %>% remove_rownames()
```




For every fusion set, find parent and fusion marker genes. Store outputs

```{r}
source('functions/find_marker_genes_ppf_lowhi.R')

#bc_markers <- find_marker_genes_ppf_lowhi(obj_trio_min, fusepairs, n_genes=100, min.pct=0.8)

#write.table(bc_markers, file=paste0('outdata/de_clonal_parent_progeny_markers/',cellline,'_parent_progeny_barcode_markers_nf'), quote=FALSE, sep='\t', col.names = TRUE)

bc_markers <- read.csv(paste0('outdata/de_clonal_parent_progeny_markers/',cellline,'_parent_progeny_barcode_markers_nf'),sep='\t') %>% remove_rownames
```



```{r}
source('functions/plot_barcode_marker_genes.R')
source('functions/plot_cluster_marker_genes.R')

scale = 1

bc_plots <- plot_barcode_marker_genes(obj_trio_min,
                                      fusepairs,
                                      include_fusion_markers = TRUE,
                                      include_control_markers = FALSE,
                                      heatmap_color_low = 'grey99',
                                      heatmap_color_mid = 'grey85',
                                      heatmap_color_hi = 'grey20',
                                      n_genes = 15,
                                      markergenes_df = bc_markers,
                                      max_cell_per_clone_plot = 100,
                                      minmax_scale = TRUE,
                                      scale_exp_min = -scale,
                                      scale_exp_max = scale,
                                      scale_exp_min_sc = -scale,
                                      scale_exp_max_sc = scale,
                                      min_counts = 20,
                                      title_size = 10,
                                      xlabel_size = 8,
                                      ylabel_size = 6
                                      )
# #
cl_plots <- plot_cluster_marker_genes(obj_trio_min,
                                      fusepairs,
                                      markers_full = markers_full,
                                      n_genes = 6,
                                      max_cell_per_clone_plot = 100,
                                      minmax_scale = TRUE,
                                      scale_exp_min = -scale,
                                      scale_exp_max = scale,
                                      scale_exp_min_sc = -scale,
                                      scale_exp_max_sc = scale,
                                      min_counts = 20,
                                      title_size = 10,
                                      xlabel_size = 8,
                                      ylabel_size = 6,
                                      heatmap_color_low = 'grey99',
                                      heatmap_color_mid = 'grey85',
                                      heatmap_color_hi = 'grey20'
                                      )



bc_plots[[1]][1]
bc_plots[[3]][1]
```

```{r}
# save single-cell heatmaps for barcode markers
for(i in 1:nrow(fusepairs)){
 fig <-  bc_plots[[1]][i][[1]]
 SaveFigure(fig, paste0(cellline,'_',fusepairs$barpairid[i],'_barcode_dot_summary'), width=2, height=4.3, res=600)
}

# save single-cell heatmaps for cluster markers
for(i in 1:nrow(fusepairs)){
 fig <-  cl_plots[[1]][i][[1]]
 SaveFigure(fig, paste0(cellline,'_',fusepairs$barpairid[i],'_cluster_dot_summary'), width=2, height=4.3, res=600)
}


# save single-cell heatmaps for barcode markers
for(i in 1:nrow(fusepairs)){
 fig <-  bc_plots[[3]][i][[1]]
 SaveFigure(fig, paste0(cellline,'_',fusepairs$barpairid[i],'_barcode_heatmap'), width=6, height=4.3, res=600)
}

# save single-cell heatmaps for cluster markers
for(i in 1:nrow(fusepairs)){
 fig <-  cl_plots[[2]][i][[1]]
 SaveFigure(fig, paste0(cellline,'_',fusepairs$barpairid[i],'_cluster_heatmap'), width=6, height=4.3, res=600)
}
```


```{r save cluster plots}
library(cowplot)

leg <- get_legend(cl_plots[[3]])

pl <- cl_plots[[1]]
iter <- 1+nrow(fusepairs)
pl[[iter]] <- leg
               

pc <- plot_grid(plotlist = pl, ncol = 7)


if(cellline=='h1806'){
  SaveFigure(pc, paste0(cellline,'_parent_progeny_cluster_markers_full_page'), width=13, height=17, res=600)
}

if(cellline=='mb231'){
  SaveFigure(pc, paste0(cellline,'_parent_progeny_cluster_markers_full_page'), width=13, height=17/2, res=600)
}
```

```{r save lineage plots}
leg <- get_legend(bc_plots[[2]])
hm_leg <- get_legend(bc_plots[[4]])

pl <- bc_plots[[1]]
iter <- 1+nrow(fusepairs)
pl[[iter]] <- leg
               

pb <- plot_grid(plotlist = pl, ncol = 7)

if(cellline=='h1806'){
  SaveFigure(pb, paste0(cellline,'_parent_progeny_barcode_markers_full_page'), width=13, height=17, res=600)
}

if(cellline=='mb231'){
  SaveFigure(pb, paste0(cellline,'_parent_progeny_barcode_markers_full_page'), width=13, height=17/2, res=600)
}
```


```{r}
if(cellline=='h1806'){
  fusepairs <- fusepairs %>% mutate(group = 1:n())
  
  focus_sets <- fusepairs #%>% filter(group %in% c(2,6,8,18,20,22)) 
  
  sc_clust <- cl_plots[[2]]
  sc_clone <- bc_plots[[3]]
  bulk_clust <- cl_plots[[1]]
  bulk_clone <- bc_plots[[1]]
  
  for(i in 1:nrow(fusepairs)){
  sc_clust_plot <- sc_clust[focus_sets$group[i]]
  sc_clone_plot <- sc_clone[focus_sets$group[i]]
  bulk_clust_plot <- bulk_clust[focus_sets$group[i]]
  bulk_clone_plot <- bulk_clone[focus_sets$group[i]]
  
  bulk_combined <-  plot_grid(plotlist = c(bulk_clust_plot, bulk_clone_plot), ncol=2)
  
  SaveFigure(bulk_combined, paste0(cellline,'_',focus_sets$barpairid[i],'_dot_plots_combined'), width=3.5, height=4.3, res=600)
  }
  
  SaveFigure(plot_grid(hm_leg), 'legend_hm', width=1, height=4.3, res=600)
  SaveFigure(plot_grid(leg), 'legend', width=1, height=4.3, res=600)
}
```


```{r save umaps}
#source('functions/ParentProgenyUMAPs_DE_functions.R')
fig_path <- "./plots/parent_progeny_barcode_markers_all/"

for(gbar in unique(fusepairs$gbarid)){
  plot_ppf_trios(obj=h1806, 
                 subset_obj = obj_trio_min, 
                 pick_clone=gbar, 
                 legend.pos = 'top', 
                 legend.title = '')
}

for(rbar in unique(fusepairs$rbarid)){
  plot_ppf_trios(obj=h1806, 
                 subset_obj = obj_trio_min, 
                 pick_clone=rbar, 
                 save_fig = T,
                 legend.pos = 'top', 
                 legend.title = '')
}
```
